---
title: "Predicting Whether a Kickstarter Project Will Succeed Or Fail"
author: "Austin Martinez, Cameron Morefield, and Derek McFate"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This report will utilize various factors from a publicly provided Kickstarter data set to determine the likelihood that a project will success or fail.
<br/>
# Data Gathering and cleaning
The data set used contains 378,661 entries, and each entry has 15 variables. There were 3797 entries that had N/A values somewhere in the row, which we decided to remove from the data set. The data set also included entries that had states other than 'Success' or 'Failure', so those entries were either condensed into the two states or dropped completely.
Once the data was cleaned, it was split into training and test data.
```{r}
set.seed(123)
library(cluster)
library(e1071)
dat = read.csv("https://raw.githubusercontent.com/AMartinez2/CST383-Project/master/ks-projects-201801.csv")
dat = dat[complete.cases(dat),]
levels(dat$state)[which(levels(dat$state)=="canceled")] = "failed"
dat = dat[!(dat$state=="undefined" | dat$state=="suspended" | dat$state=="live"),]
# split our data into testing and training data
tr_rows = sample(nrow(dat), 0.8 * nrow(dat))
tr_dat = dat[tr_rows,]
te_dat = dat[-tr_rows,]
```

# Initial Exploration and visualization

Perhaps the goal for the project influences how people will look at project which would affect if they choose to contribute. If we look at the density of goal for projects that passed (in green) and failed (in red), we can see that values align well within the same curve. The goal of the project does not seem to affect wether it passes or fails. 

```{r}
pass = dat[dat$state == "successful",]
fail = dat[dat$state == "failed",]
plot(density(pass$goal), xlim=c(0,100000), col='green', main='Double Density of Goal')
lines(density(fail$goal), col='red')
```

######### We might want to remove this ##############
```{r}
par(mar=c(3,6,1,4))
pass = dat[dat$state == "successful",]
fail = dat[dat$state == "failed",]
test2 <- rbind(table(pass$main_category),table(fail$main_category))
barplot(test2,beside=T, col=c('green','red'), horiz=TRUE, las=1)
```

```{r}

```

# First model (Naive Bayes)
```{r}
fit = naiveBayes(state ~ category, data=tr_dat)
predicts = predict(fit, newdata=te_dat)
conf_mtx = table(predicts, te_dat$state)
conf_mtx[c('failed','successful'),c('failed','successful')]
mean(predicts == te_dat$state)
```

```{r}
library(rpart)

fit = rpart(state ~ goal, data = tr_dat)
predicted = predict(fit, te_dat, type="class")
actual = te_dat$state
table(actual, predicted)
mean(actual == predicted)
```

# Second Model 

```{r}
fit = glm(state ~ main_category + goal, data=tr_dat, family=binomial)
y = predict(fit, newdata=te_dat, type="response")
predicts = ifelse(y > 0.5, "successful", "failed")
actuals = te_dat$state
conf_mtx = table(predicts, actuals)
conf_mtx
mean(predicts == actuals)
```

```{r}

```

# Conclusion

